Boson-Sampling with Single Photons: A Roadmap to Post-Classical Computation

### 2016-09-21 backup.tex
almost there

@@

@@

@@

It was proven by Reck \emph{et al.} \cite{bib:18} that an arbitrary $m$-mode linear optical network can be efficiently constructed using $O(m^2)$ linear optical elements (i.e. beamsplitters and phase-shifters). This can be reduced to $O(mn)$ in the case of boson-sampling, where there are $n$ photons in $m$ modes. Although `efficient', this scaling implies that constructing a network with hundreds of modes with discrete elements, one probably needs several room-sized optical tables, inevitably suffering from severe air turbulence and mechanical instability. For perfect quantum interference, photons should arrive at the points of interference simultaneously and be perfectly mode-matched (to be discussed in more detail in Sec.~\ref{fidelity_of_devices}), which means all of the thousands of optical elements must be carefully simultaneously aligned. More integrated and stable alternatives are of pressing importance. In fact, all initial demonstrations of boson-sampling were based on integrated optical circuits including, waveguides based on direct laser writing techniques \cite{bib:4, bib:5, bib:6} and fused fibre couplers \cite{bib:10}.

>>>>

Boson-sampling is a process in which $n$ indistinguishable photons are injected into an $m$-mode linear optical interferometer, characterised by linear optics transformation $U$ (i.e beamsplitters and phase-shifters), and then detected at output ports using photodetectors, as shown in Fig.~\ref{fig:model} \cite{bib:2}. $U$ evolves the photonic creation operators according to,
\begin{align}
\hat{U} \hat{a}^\dag_i \hat{U}^\dag \to \sum_{j=1}^m U_{i,j} \hat{a}^\dag_j,
\end{align}
where $\hat{a}^\dag_i$ is the photon creation operator for the $i$th mode.

Choosing the mode occupation-number configurations as the basis states, the input state can be written as a list of occupation numbers \mbox{$\ket{S}=\ket{s_1,\dots,s_m}$}, where $s_i$ is the number of photons in the $i$th input mode. As photon-number is preserved via linear optics transformations, the output state can be denoted in the same basis, but a superposition of all different output configurations,
\begin{align}
\ket\varphi = \sum_{i=1}^{|\Phi|}\alpha_i \ket{T_i},
\end{align}
where \mbox{$\ket{T_i} = \ket{t_1,\dots,t_m}$}, and with each basis state satisfying
\begin{align}
\sum_{i=1}^m s_i = \sum_{i=1}^m t_i = n \,\,\forall\,\, S,T.
\end{align}
We define $\Phi$ as the set of all photon-number configuration basis vectors. It is not hard to prove that the total number of elements in $\Phi$ is,
\begin{align}
|\Phi|=\binom{m+n-1}{n}.
\end{align}
The amplitude associated with each output basis is given by \cite{bib:2, bib:12},
\begin{align}
\bra{T_i} \Phi(\hat{U}) \ket{S} = \frac{\mathrm{Per}(U_{S,T_i})}{\sqrt{s_1!\dots s_m!t_1!\dots t_m!}},
\end{align}
where $\Phi(U)$ is a homomorphism of $U$, and $U_{S,T_i}$ is an $n\times n$ submatrix of $U$, obtained by taking $s_i$ copies of the $i$th row of $U$, and $t_i$ copies of the $i$th column of $U$. It has been proven that \cite{bib:2, bib:13, bib:14} one will observe a collision (i.e. two or more photons in the same mode) with probability bounded away from 0 only if the number of modes scales as at least $m=O(n^2)$ -- the so-called `bosonic birthday paradox' \cite{???}. In this instance, it suffices to use non-photon-number-resolving detectors, substantially simplifying technological requirements. Thus, throughout this paper we only discuss the `collision-free' situation, i.e. $s_i,t_j=\{0,1\}$.

Consider the practical situation, where we have an imperfect photonic simulator to implement boson-sampling, whose net per-photon efficiency is $\eta$, combining the single-photon generation rate, circuit transmission efficiency, and coupling and detection efficiency (in a linear optics network, all losses can be commuted to the front or back of the circuit, provided all losses are uniform across all modes, allowing us to merge all the individual efficiencies into a single overall efficiency factor). The time required to obtain an $n$-photon coincidence event is,
\begin{align}
\tau_q = \frac{1}{\eta^n R}
\end{align}
where $R$ is the effective repetition rate of the quantum boson-sampler, restricted by the choice of single photon source (SPS), the experimental architecture, and the dead-time of the detectors.

@@

How can a classical computer simulate this bosonic process, and sample from its exact or approximate probability distribution? A na{\"i}ve brute-force method is to calculate all $|\Phi|$ permanents of the set of submatrices $U_{S,T_i}$, thereby determining the full probability distribution, and then sample from this known distribution. \textbf{However, neither does the boson-sampler know even the approximate probability distribution before enough samples are obtained} \comment{What does that sentence mean?}. Smarter classical algorithms exist to generate correct samples without knowing the entire underlying distribution. For example, using the acceptance-rejection method \cite{bib:15} one can sample an output configuration $\ket{T_i}$ uniformly from the configuration space and a number $u$ from the uniform distribution over $\{0,1\}$, then calculate the permanents of the respective submatrices and compare them to $u$; If \mbox{$u<|\mathrm{Per}(U_{S,T_i})|^2$}, we accept the sample $\ket{T_i}$; If not, it is rejected, and we repeat. Such methods could be successful with probability $O(1)$. One sample only requires calculating one permanent, significantly better than the brute-force method of calculating the entire distribution.

The permanent is a matrix function originating from the permutation symmetry of bosons, similar to the role of determinants for fermions. Na{\"i}ve calculation of the permanent of an $n\times n$ matrix requires $n!$ operations. However, a more efficient classical algorithm developed by Ryser \cite{bib:16} in 1963 requires $O(n^2 2^{n+1})$ operations, which is nonetheless exponential. No general polynomial-time algorithms have been described, in contrast to the clever Gaussian elimination method for calculate matrix determinants. In fact, calculating the permanent was proved to be \#P-complete in 1979 by Valiant \cite{bib:17}. A more intuitive understanding is that the permanent corresponds to counting the number of different matching ways between input state and output state through the interferometer, much like adding Feynman paths \cite{} to determine quantum amplitudes.

Consequently, for a classical simulator, the time required to acquire a sample using this approach is,
\begin{align}
\tau_c \approx \frac{n^2 2^{n+1}}{A},
\end{align}
where $A$ is number of floating-point operations per second (FLOPS) of the classical computer.

@@

Comparing the requisite times, we plot the critical photon-number as a function of efficiency $\eta$ for the quantum boson-sampler to beat its classical counterpart, as shown in Fig.~\ref{fig:curves_1}(a). The right part of each line is the area of quantum supremacy. For a given single-photon efficiency, with increasing $n$, the quantum simulator's superiority becomes more prominent. Thus, scaling up the size of boson-sampling is essential to achieving quantum supremacy. But there is an upper limit. Taking \mbox{$\eta=0.4$} as an example, it is impossible to reach quantum supremacy for a classical computer of petaFLOPS by only enlarging the number of single photons. A more convenient approach is to increase the efficiency of the quantum system. As the efficiency increases, the minimum photon-number required to achieve quantum supremacy decreases, simplifying experimental resource requirements.

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{curves_1}
\caption{The quantum and classical supremacy of (spatially-encoded) boson-sampling is divided by the red lines, where the top-right region corresponds to quantum supremacy and vice versa. The efficiency denoted on the horizontal axis is the overall system efficiency. (a) $R=76$MHz. The numbers 3, 6, etc. in the lines refer to $10^3\mathrm{FLOPS}$, $10^6\mathrm{FLOPS}$ etc. of the classical simulator. (b) $A=10^9\mathrm{Hz}$. The numbers 2, 3, etc. in the lines denote photon-number.} \label{fig:curves_1}
\end{figure}

The repetition rate $R$ is another important factor affecting the performance of quantum simulators that we must consider in practical experiments. If $\eta^n$ is taken as the post-selection success probability of a single trial, then \mbox{$\eta^n R$} indicates how many successful runs we can implement per second. For example, to beat a gigaFLOPS classical computer, four-photon boson-sampling with $R=76\mathrm{MHz}$ ($R$ vs $A$) is sufficient if the total efficiency is \mbox{$\eta=0.4$}. As shown in Fig.~\ref{fig:curves_1}(b), however, \mbox{$\eta=0.2$} and \mbox{$n=2$} is sufficient if the repetition rate increases to 1GHz. In other words, by increasing the repetition rate, the efficiency and photon-number requirements are relaxed. However, we cannot increase the repetition rate indefinitely. It is restricted by the pulse rate of the SPS, experimental architecture, and dead-time of the detectors. In the next section we discuss the impact of the choice of experimental architecture.

Here we must point out two things. First, Ryser's algorithm is actually not faster than using the original formula to calculate the matrix permanent when photon number is small (\mbox{$n<6\sim 7$}). Thus the above conditions may be not tight. For example, the efficiency must be larger than $\sim 0.6$ if two- (four)-photon boson-sampling with a repetition rate of 1GHz (76MHz) is expected to beat a gigaFLOPS classical computer.

Why then should we nonetheless scale up to more photons (\mbox{$n=20\sim 30$}), even though boson-sampling with relatively few photons could beat current prevailing personal computers in realistic experimental regimes? One reason is that no finite experiment can make a decisive conclusion about the ECT, since it is a conjecture about the asymptotic limit. In the field of computational complexity, the significance of boson-sampling rests largely on its exponentially faster sampling speed than that of classical simulators, rather than certain times of specific triumphs in the competitions between quantum and classical computers. Thus, through upscaling we can make stronger statements about asymptotic behaviour (albeit never a completely conclusive one). Another cause is that people have not physically constructed relatively large multi-photon interferometers yet. The theory of boson-sampling still faces the risk of failure and quantum devices may break down with more photons. Finally, although no practical uses for boson-sampling have been described, it is entirely plausible that useful applications may be found in the future. Were that the case, scaling up boson-sampling experiments would similarly scale up the size of the problems it could solve.

@@

It was proven by Reck \emph{et al.} \cite{bib:18} that an arbitrary $m$-mode linear optical network can be efficiently constructed using $O(m^2)$ linear optical elements (i.e. beamsplitters and phase-shifters). This can be reduced to $O(mn)$ in the case of boson-sampling, where there are $n$ photons in $m$ modes. Although `efficient', this scaling implies that constructing a network with hundreds of modes with discrete elements, one probably needs several room-sized optical tables, inevitably suffering from severe air turbulence and mechanical instability. For perfect quantum interference, photons should arrive at the points of interference simultaneously and be perfectly mode-matched (to be discussed in more detail in Sec.~\ref{sec:fidelity_of_devices}), which means all of the thousands of optical elements must be carefully simultaneously aligned. More integrated and stable alternatives are of pressing importance. In fact, all initial demonstrations of boson-sampling were based on integrated optical circuits including, waveguides based on direct laser writing techniques \cite{bib:4, bib:5, bib:6} and fused fibre couplers \cite{bib:10}.
